Accurate evaluation of dislocation tractions

### 2021-03-10 src/wip/DanielCelisGarza/Accurate_evaluation_of_dislocation_tractions/paper/main.tex
Emphasis, minor typos.

Coupling discrete dislocation dynamics (DDD) to the finite element method (FEM) allows micromechanical tests to be simulated with explicit dislocation interactions \cite{Groh2009}. Virtual experiments are essential for interpreting experimental data and relating the measured mechanical response to dislocation mechanisms \cite{0965-0393-6-6-007,doi:10.1080/14786430500341250,tarleton2015discrete,YU2018}. One of the most popular and simplest methods for coupling DDD and FEM is the superposition method \cite{superposition_scheme0,superposition_scheme1,superposition_scheme2}, which works by decomposing the problem into separate DDD and FE problems as illustrated in \cref{f:superposition_scheme}.
%
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{fem_ddd.pdf}
    \caption[Superposition Model for DDD-FEM coupling.]{The superposition used to couple DDD and FEM. The volume $V$ is bounded by a surface $S = S_{T} \cup S_{U}$ and contains a dislocation ensemble and is subjected to tractions $\vec{T}$ on $S_{T}$ and $\vec{u}$ on $S_{u}$. First, the traction, $\vec{\tilde{T}}$, and displacement, $\vec{\tilde{u}}$, fields due to the dislocations in the infinite domain (DDD) are evaluated on the boundaries $S_{T}$ and $S_{U}$ respectively. Then an elastic boundary value problem can be solved with FEM to calculate the corrective elastic fields required to satisfy the boundary conditions $\vec{\hat{T}} = \vec{T} - \vec{\tilde{T}}$ and $\vec{\hat{u}} = \vec{U} - \vec{\tilde{U}}$.}
    \label{f:superposition_scheme}
\end{figure}

Specifically, a linear-elastic solid $V$ bounded by a surface $S$ is subjected to traction boundary conditions, $\vec{T}$, on $S_{T}$ and displacement boundary conditions, $\vec{U}$, on $S_{U}$. The ($\tilde{~}$) fields are those generated by the dislocations in an infinite solid and in our case are obtained by evaluating analytic solutions in a DDD simulation \cite{Cai2006}. Formally, the dislocation field satisfies,
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\tilde{\sigma}} = \vec{0}                \\
        \tns{\tilde{\sigma}} = \tns{C} : \tns{\tilde{\varepsilon}} \\
        \tns{\tilde{\varepsilon}} = \frac{1}{2} \left( \nabla \vec{\tilde{u}} + (\nabla \vec{\tilde{u}})^T \right)
    \end{array}
    \right\} \quad \textrm{in V}                                                   \\
    %
     & \tilde{\tns{\sigma}} \cdot \vec{n} = \tilde{\vec{T}} \quad \textrm{on } S_T \\
    %
     & \left.
    \begin{array}{l}
        \vec{\tilde{u}} = U, \quad t>0 \\
        \vec{\tilde{u}} = \vec{0}, \quad t=0
        \label{eq:utildebc}
    \end{array}
    \right\} \quad \textrm{on } S_U \,.
\end{align}
%
As the dislocation fields do not vanish on $S$, the dislocations load the volume by generating tractions, $\tilde{\vec{T}}$, on $S_T$ and displacements, $\vec{\tilde{u}}$, on $S_U$. This additional loading deforms $V$, generating an additional ``image'' stress which the dislocations then feel. Therefore corrective ($\hat{~}$) fields must be superimposed to satisfy the desired boundary conditions. The corrective field which accounts for both the applied and image stress is obtained numerically by solving the elastic boundary value problem,
%
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\hat{\sigma}} = \vec{0}              \\
        \tns{\hat{\sigma}} = \tns{C} : \tns{\hat{\varepsilon}} \\
        \tns{\hat{\varepsilon}} = \frac{1}{2} \left( \nabla \vec{\hat{u}} + (\nabla \vec{\hat{u}})^{T} \right)
    \end{array}\right\} \quad \textrm{in V}                                                  \\
    %
     & \hat{\tns{\sigma}}\cdot\vec{n} = \vec{T} - \tilde{\vec{T}} \quad \textrm{on } S_T \label{eq:thatbc} \\
    %
     & \vec{\hat{u}} = \vec{U} - \vec{\tilde{u}} \quad \textrm{on } S_U \,. \label{eq:uhatbc}
\end{align}
%
Once the solutions to both problems are known, their superposition solves the desired mixed boundary value problem,
%
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\sigma} = \nabla \cdot \left( \tns{\hat{\sigma}} + \tns{\tilde{\sigma}} \right) = \vec{0}                                                             \\
        \tns{\sigma} = \tns{\hat{\sigma}} + \tns{\tilde{\sigma}} = \tns{C} : \left( \tns{\hat{\varepsilon}} + \tns{\tilde{\varepsilon}} \right) = \tns{C} : {\tns{\varepsilon}} \\
        \tns{\varepsilon} = \tns{\hat{\varepsilon}} + \tns{\tilde{\varepsilon}} = \frac{1}{2} \left( \nabla ( \vec{\hat{u}} + \vec{\tilde{u}} ) + \left[ \nabla ( \vec{\hat{u}} + \vec{\tilde{u}} ) \right]^T \right) = \frac{1}{2} \left( \nabla \vec{u} + ( \nabla \vec{u} )^{T} \right)
    \end{array}
    \right\}\quad \textrm{in V}                                    \\
    %
     & \tns{\sigma} \cdot \vec{n} = \vec{T} \quad \textrm{on } S_T \\
    %
     & \left.
    \begin{array}{l}
        \vec{u} = \vec{U}, \quad t>0 \\
        \vec{u} = \vec{0}, \quad t=0
    \end{array}
    \right\} \quad \textrm{on } S_U
    \label{eq:ubc}
\end{align}

@@

The units in all our examples are normalised to lattice parameter, $a$ is the dislocation core radius for the non-singular formulation discussed in \cite{Cai2006}, and $b \equiv \lVert \vec{b} \rVert$.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{head_vs_fe_tractions_setup_diagram.pdf}
    \caption{Dislocation parallel to a surface described by the line $x = 0$, where the $(x,\,y)$ dislocation coordinates are $(a,\,c)$ and the dislocation line going in the positive $z$-direction. (a) describes the box we used for our comparison, a  $40 \times 40 \times 40$ element cubic box with side lengths equal to $2000$ with $\vec{T} = \vec{0}$, traction boundary conditions only on the nodes along $x = 0$, and $\vec{U} = \vec{0}$, displacement conditions everywhere else. (b) is the 2D view with the dislocation coordinates $(a,~c)$, as well as the two edge Burgers vectors in \cite{head1953edge}, $\vec{b}_\textrm{e1} \equiv \vec{b} = [1\, 0\, 0]$, $\vec{b}_\textrm{e2} \equiv \vec{b} = [0\, 1\, 0]$ and the screw Burgers vector $\vec{b}_\textrm{s} \equiv \vec{b} = \vec{l} = [0\, 0\, 1] $ in \cite[p.~59,~64]{hirth1983theory}.}
    \label{f:headvstractionfem}
\end{figure}
The slices we took for our contour plots in \cref{s:results} are on the middle plane of the domain at $z = 1000\, $.

@@

    \subfloat[Segment starts 201 dislocation core radii away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.475\linewidth]{perp_e_xz=1005_leg.pdf}}
    }
    ~
    \subfloat[Segment starts 20001 dislocation core radii away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.47\linewidth]{perp_e_xz=100005_leg.pdf}}
    }
    \caption{Log-log plot of the relative error as a function of dislocation segment length for a perpendicular edge dislocation (\cref{f:gauss_quad_test}a). $x^{1}_{z}$ is the $z$-coordinate of node $\mathbf{x_1}$. $\vec{b} = [1\, 0\, 0]$, line direction, $\vec{l} = [0\, 0\, 1]$, and dislocation core radius, $a = 5$, the surface element's normal and size are, $\vec{n} = [0\, 0\, 1]$, $L = 1000$, respectively. $Q$ is the number of quadrature points per dimension.}
    \label{f:rel_err_perp_edge}
\end{figure}
It is worth noting however that the relative errors for small numbers of quadrature points don't really start decreasing until relatively large distances. And when close to a surface, these can be quite large even under hihgly symmetric circumstances.

>>>>

Coupling discrete dislocation dynamics (DDD) to the finite element method (FEM) allows micromechanical tests to be simulated with explicit dislocation interactions \cite{Groh2009}. Virtual experiments are essential for interpreting experimental data and relating the measured mechanical response to dislocation mechanisms \cite{0965-0393-6-6-007,doi:10.1080/14786430500341250,tarleton2015discrete,YU2018}. One of the most popular and simplest methods for coupling DDD and FEM is the superposition method \cite{superposition_scheme0,superposition_scheme1,superposition_scheme2}, which works by decomposing the problem into separate DDD and FE problems as illustrated in \cref{f:superposition_scheme}.
%
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{fem_ddd.pdf}
    \caption[Superposition Model for DDD-FEM coupling.]{The superposition used to couple DDD and FEM. The volume $V$ is bounded by a surface $S = S_{T} \cup S_{U}$ and contains a dislocation ensemble and is subjected to tractions $\vec{T}$ on $S_{T}$ and $\vec{U}$ on $S_{u}$. First, the traction, $\vec{\tilde{T}}$, and displacement, $\vec{\tilde{u}}$, fields due to the dislocations in the infinite domain (DDD) are evaluated on the boundaries $S_{T}$ and $S_{U}$ respectively. Then an elastic boundary value problem can be solved with FEM to calculate the corrective elastic fields required to satisfy the boundary conditions $\vec{\hat{T}} = \vec{T} - \vec{\tilde{T}}$ and $\vec{\hat{u}} = \vec{U} - \vec{\tilde{u}}$.}
    \label{f:superposition_scheme}
\end{figure}

Specifically, a linear-elastic solid $V$ bounded by a surface $S$ is subjected to traction boundary conditions, $\vec{T}$, on $S_{T}$ and displacement boundary conditions, $\vec{U}$, on $S_{U}$. The ($\tilde{~}$) fields are those generated by the dislocations in an infinite solid and in our case are obtained by evaluating analytic solutions in a DDD simulation \cite{Cai2006}. Formally, the dislocation field satisfies,
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\tilde{\sigma}} = \vec{0}                \\
        \tns{\tilde{\sigma}} = \tns{C} : \tns{\tilde{\varepsilon}} \\
        \tns{\tilde{\varepsilon}} = \frac{1}{2} \left( \nabla \vec{\tilde{u}} + (\nabla \vec{\tilde{u}})^T \right)
    \end{array}
    \right\} \quad \textrm{in V}                                                   \\
    %
     & \tilde{\tns{\sigma}} \cdot \vec{n} = \tilde{\vec{T}} \quad \textrm{on } S_T \\
    %
     & \left.
    \begin{array}{l}
        \vec{\tilde{u}} = \vec{U}, \quad t>0 \\
        \vec{\tilde{u}} = \vec{0}, \quad t=0
        \label{eq:utildebc}
    \end{array}
    \right\} \quad \textrm{on } S_U \,.
\end{align}
%
As the dislocation fields do not vanish on $S$, the dislocations load the volume by generating tractions, $\tilde{\vec{T}}$, on $S_T$ and displacements, $\vec{\tilde{u}}$, on $S_U$. This additional loading deforms $V$, generating an additional ``image'' stress which the dislocations then feel. Therefore corrective ($\hat{~}$) fields must be superimposed to satisfy the desired boundary conditions. The corrective field which accounts for both the applied and image stress is obtained numerically by solving the elastic boundary value problem,
%
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\hat{\sigma}} = \vec{0}              \\
        \tns{\hat{\sigma}} = \tns{C} : \tns{\hat{\varepsilon}} \\
        \tns{\hat{\varepsilon}} = \frac{1}{2} \left( \nabla \vec{\hat{u}} + (\nabla \vec{\hat{u}})^{T} \right)
    \end{array}\right\} \quad \textrm{in V}                                                  \\
    %
     & \hat{\tns{\sigma}}\cdot\vec{n} = \vec{T} - \tilde{\vec{T}} \quad \textrm{on } S_T \label{eq:thatbc} \\
    %
     & \vec{\hat{u}} = \vec{U} - \vec{\tilde{u}} \quad \textrm{on } S_U \,. \label{eq:uhatbc}
\end{align}
%
Once the solutions to both problems are known, their superposition solves the desired mixed boundary value problem,
%
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\sigma} = \nabla \cdot \left( \tns{\hat{\sigma}} + \tns{\tilde{\sigma}} \right) = \vec{0}                                                             \\
        \tns{\sigma} = \tns{\hat{\sigma}} + \tns{\tilde{\sigma}} = \tns{C} : \left( \tns{\hat{\varepsilon}} + \tns{\tilde{\varepsilon}} \right) = \tns{C} : {\tns{\varepsilon}} \\
        \tns{\varepsilon} = \tns{\hat{\varepsilon}} + \tns{\tilde{\varepsilon}} = \frac{1}{2} \left( \nabla ( \vec{\hat{u}} + \vec{\tilde{u}} ) + \left[ \nabla ( \vec{\hat{u}} + \vec{\tilde{u}} ) \right]^T \right) = \frac{1}{2} \left( \nabla \vec{u} + ( \nabla \vec{u} )^{T} \right)
    \end{array}
    \right\}\quad \textrm{in V}                                    \\
    %
     & \tns{\sigma} \cdot \vec{n} = \vec{T} \quad \textrm{on } S_T \\
    %
     & \left.
    \begin{array}{l}
        \vec{u} = \vec{U}, \quad t>0 \\
        \vec{u} = \vec{0}, \quad t=0
    \end{array}
    \right\} \quad \textrm{on } S_U
    \label{eq:ubc}
\end{align}

@@

The units in all our examples are normalised to lattice parameter, $a$ is the dislocation core radius for the non-singular formulation discussed in \cite{Cai2006}, and $b \equiv \lVert \vec{b} \rVert$.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{head_vs_fe_tractions_setup_diagram.pdf}
    \caption{Dislocation parallel to a surface described by the line $x = 0$, where the $(x,\,y)$ dislocation coordinates are $(a,\,c)$ and the dislocation line going in the positive $z$-direction. (a) describes the box we used for our comparison, a  $40 \times 40 \times 40$ element cubic box with side lengths equal to $2000$ with $\vec{T} = \vec{0}$, traction boundary conditions only on the nodes along $x = 0$, and $\vec{U} = \vec{0}$, displacement conditions everywhere else. (b) is the 2D view with the dislocation coordinates $(a,~c)$, as well as the two edge Burgers vectors in \cite{head1953edge}, $\vec{b}_\textrm{e1} \equiv \vec{b} = [1\, 0\, 0]$, $\vec{b}_\textrm{e2} \equiv \vec{b} = [0\, 1\, 0]$ and the screw Burgers vector $\vec{b}_\textrm{s} \equiv \vec{b} = \vec{l} = [0\, 0\, 1] $ in \cite[p.~59,~64]{hirth1983theory}.}
    \label{f:headvstractionfem}
\end{figure}
The slices we took for our contour plots in \cref{s:results} are on the middle plane of the domain at $z = 1000$.

@@

    \subfloat[Segment starts 201 dislocation core radii away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.475\linewidth]{perp_e_xz=1005_leg.pdf}}
    }
    ~
    \subfloat[Segment starts 20001 dislocation core radii away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.47\linewidth]{perp_e_xz=100005_leg.pdf}}
    }
    \caption{Log-log plot of the relative error as a function of dislocation segment length for a perpendicular edge dislocation (\cref{f:gauss_quad_test}a). $x^{1}_{z}$ is the $z$-coordinate of node $\mathbf{x_1}$. $\vec{b} = [1\, 0\, 0]$, line direction, $\vec{l} = [0\, 0\, 1]$, and dislocation core radius, $a = 5$, the surface element's normal and size are, $\vec{n} = [0\, 0\, 1]$, $L = 1000$, respectively. $Q$ is the number of quadrature points per dimension.}
    \label{f:rel_err_perp_edge}
\end{figure}
It is worth noting however that the relative errors for small numbers of quadrature points don't really start decreasing until relatively large distances. And when close to a surface, these can be quite large even under highly symmetric circumstances.
